---
title: "Data Visualization Utilities"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Data Visualization Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.width = 7,
  fig.height = 5,
  fig.align = 'center',
  warning = FALSE,
  message = FALSE
)
```

## Setup

The demonstration of `yawp` usage in this vignette depends on functions and data sets from `ggplot2`, `patchwork`, `dplyr`, and `tidyr`:

```{r setup}
library(yawp)
library(ggplot2)
library(patchwork)
library(tidyr)
library(dplyr)
```

## Themes

### `theme_mathbook()`

```{r}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point(col = "white") +
  geom_smooth(method = "lm", col = "firebrick", se = FALSE) +
  facet_wrap(~ trans, ncol = 5) +
  labs(x = "Displacement (Liters)", y = "Highway MPG") +
  theme_mathbook()
```

## Zoom in on a `ggplot2` plot

`gg_zoom()` takes a `ggplot2` plot and "zooms in" on a user-specified subset of the data. The filtering operation is defined in the "zoom_cmd" argument, which accepts the same syntax as the `filter()` function from the `dplyr` package:

```{r}
p <-
  population %>%  
  filter(country %in% c("Finland","Norway","Sweden", "Denmark")) %>%
  ggplot(aes(year, population)) +
  geom_line(aes(group = country, col = country) , lwd = 2) +
  labs(x = "Year", y = "Population") +
  theme_minimal() +
  theme(legend.position = "bottom", legend.title = element_blank())

p %>%
  gg_zoom(., zoom_cmd = year >=2005 & year <= 2010, draw_box = FALSE)
  
```

The function includes an option to label the points with another column:

```{r}
p <-
  msleep %>%
  filter(!grepl("elephant", name)) %>%
  ggplot(aes(brainwt, bodywt)) +
  geom_point() +
  labs(x = "Brain weight (kilograms)", y = "Body weight (kilograms)") +
  theme_minimal() +
  theme(legend.position = "bottom")

p %>%
  gg_zoom(., zoom_cmd = brainwt < 0.0005, draw_box = FALSE, to_label = TRUE, label = name)
```

`gg_zoom()` works with discrete axes as well:

```{r}
p <-
  msleep %>%
  ggplot(aes(vore, sleep_total)) +
  geom_violin() +
  labs(x = "Vore", y = "Daily sleep total (hours)") +
  theme_minimal()

gg_zoom(p, zoom_cmd = vore %in% c("carni", "herbi"), draw_box = FALSE)
```

The returned value is a `patchwork` object, which means that additional formatting from the [`patchwork` package](https://CRAN.R-project.org/package=patchwork) can be applied:

```{r}
gg_zoom(p, zoom_cmd = vore %in% c("carni", "herbi"), draw_box = FALSE) +
  plot_layout(widths = c(3,1)) +
  plot_annotation(title = "Amount of Sleep by Vore")
```

```{r}
layout <- c(
  area(t = 1, l = 1, b = 2, r = 4),
  area(t = 3, l = 4, b = 4, r = 5)
)

gg_zoom(p, zoom_cmd = vore %in% c("carni", "herbi"), draw_box = FALSE) +
  plot_layout(design = layout) +
  plot_annotation(title = "Amount of Sleep by Vore")
```

