---
title: "Data Manipulation and Analysis Utilities"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Data Manipulation and Analysis Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = 'center'
)
```

## Setup

The demonstration of `yawp` usage in this vignette depends on functions and data sets from `ggplot2`, `dplyr`, and `tidyr`:

```{r setup}
library(yawp)
library(ggplot2)
library(dplyr)
library(tidyr)
theme_set(theme_minimal())
```

## Manipulate a vector

### Extract first character

The package includes `first_char()`, which is a simple function to pull the first character from a string:

```{r}
first_char("eye")
```

The function is case-sensitive:

```{r}
first_char("Eye")
```

It can also extract first characters that are not letters:

```{r}
first_char("3ye")
```

```{r}
first_char("!eye")
```

### Bound at minimum and maximum

The `bound()` function allows you to constrain a vector at maximum and minimum values:

```{r}
x <- c(-10,-5,0,2,5,10,100)

bound(x, min = 0, max = 10)
```

```{r}
x <- rnorm(1000, mean = 0.5, sd = 1)

data.frame(x = x) %>%
  ggplot(aes(x)) +
  geom_histogram()
```
```{r}
x <- bound(x, min = 0.1, max = 0.9)

data.frame(x = x) %>%
  ggplot(aes(x)) +
  geom_histogram()
```

### Find values between upper and lower limits

`betwixt()` searches for values that fall between upper and lower boundaries:

```{r}
x <- 50:60

betwixt(x, lower = 52, upper = 58)
```

By default the function returns a logical vector of indices that match the range specified, however the `value = TRUE` setting allows values to be returned instead:

```{r}
betwixt(x, lower = 52, upper = 58, value = TRUE)
```

The function can also accommodate arbitrary increments in the sequence:

```{r}
betwixt(x, lower = 52, upper = 58, value = TRUE, by = 2)
```

And it can ignore certain values as well:

```{r}
betwixt(x, lower = 52, upper = 58, value = TRUE, by = 2, ignore = 54)
```

The `betwixt()` utility can also operate on a vector of dates:

```{r}
dates <- seq(as.Date("2000/1/1"), by = "day", length.out = 365)

betwixt(dates,
        lower = as.Date("2000/1/10"),
        upper = as.Date("2000/2/10"),
        by = "day",
        ignore = as.Date("2000/1/25"),
        value = TRUE)
```

The logical index option (`value = FALSE`) can be useful in subsetting operations:

```{r}
tibble(date = dates,
       y = rnorm(365)) %>%
filter(betwixt(date, 
               lower = as.Date("2000/1/10"), 
               upper = as.Date("2000/1/17"), 
               by = "day", 
               ignore = as.Date(c("2000/1/12", "2000/1/13")),
               value = FALSE))
```

## Summarize and format

### Calculate mode

The `get_mode()` function calculates the most frequent observations:

```{r}
x <- c(1,1,2,3,3,3,3,3,4)

get_mode(x)
```

If there are ties, the function can return tied modes with `ties = TRUE`:

```{r}
x <- c(1,1,2,3,3,3,3,3,4,4,4,4,4)

get_mode(x, ties = TRUE)
```

The function can also return a "mode" for a character vector:

```{r}
x <- c("dog","dog","cat","rat","rat","rat","rat")

get_mode(x)
```


### Proportion

The `propf()` utility calculates the count and proportion of values in a vector that match a given value specified by the "level" argument. The output is formatted as string with count and percentage by default:

```{r}
dogs <- c("dog","dog","dog")
cats <- c("cat","cat")
rats <- c("rat","rat","rat","rat","rat")
animals <- c(dogs,cats,rats)

propf(animals, level = "rat")
```

The output may be customized to show proportion rather than percentage:

```{r}
propf(animals, level = "rat", percent = FALSE)
```

And the function may also accept arguments to `base::formatC()`:

```{r}
propf(animals, level = "rat", percent = FALSE, decimal.mark = ",")
```

Formatting counts and proportions can be useful when summarizing data in a table:

```{r}
starwars %>%
  select(films,sex) %>%
  unnest(films) %>%
  group_by(films) %>%
  summarise(female = propf(sex, level = "female"))
```

### Median

The `medf()` function is very similar to `propf()`, except that it calculates the median and 25th, 75th quartiles of a numeric vector:

```{r}
x <- rpois(1000, lambda = 3)

medf(x)
```

Like `propf()`, this function can take additional arguments to `base::formatC()`:

```{r}
medf(x, drop0trailing = TRUE)
```

The function can be useful for summarizing data in tables too:

```{r}
state_info <-
  tibble(state = state.name,
         region = state.division)

us_rent_income %>%
  select(NAME, variable, estimate) %>%
  spread(variable, estimate) %>%
  filter(!NAME %in% c("District of Columbia", "Puerto Rico")) %>%
  rename(state = NAME) %>%
  left_join(state_info) %>%
  group_by(region) %>%
  summarise(income = medf(income),
            rent = medf(rent))
```


### Mean and standard error

`summary_se()` summarizes data by calculating the mean, standard error and a confidence interval across groups. The function accepts bare column names for the variable to summarize ("measure_var") and the optional grouping columns:

```{r}
ToothGrowth %>%
  summary_se(len, supp, dose, .ci = 0.95)
```

```{r}
ToothGrowth %>%
  summary_se(len, supp, dose, .ci = 0.95) %>%
  mutate(dose = paste0("Dose: ", dose, " (mg/day)")) %>%
  ggplot(aes(supp,mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - ci, 
                    ymax = mean + ci),
                width = 0.2) +
  labs(x = "Vitamin C delivery method", y = "Mean length of odontoblasts (95% CI)") +
  coord_flip() +
  facet_wrap(~ dose, ncol = 1)
```

